Formal Proof of the Relative Consistency of ZF$+\neg$AC with Isabelle/ZF

### 2024-12-24 paper/main.tex
updated

\chapter*{Abstract}
We formalize the relative consistency proof of ZF$+\neg$AC using Isabelle/ZF proof assistant.
Our approach assumes the existence of a transitive countable model of ZF and uses forcing to construct a symmetric extension which is a model of ZF$+\neg$AC.
We show that the symmetric extension satisfies ZF$+\neg$AC by formalizing a relativized forcing relation based on the formalization of forcing by Gunther et al.

>>>>

\chapter*{Abstract}
This paper presents a formalization of the relative consistency of the negation of the Axiom of Choice ($\neg$AC) 
with respect to Zermelo-Fraenkel set theory (ZF) using the proof assistant Isabelle/ZF. 
We use the countable transitive model (c.t.m.) approach, assuming the existence of a c.t.m.\ of ZF, 
and apply forcing to construct a symmetric extension called the basic Cohen model, 
which serves as a model of ZF + $\neg$AC. This work builds on previous formalizations, 
such as Gunther et al.'s work on forcing, and provides the first formalization of the relative consistency of $\neg$AC with ZF.

The proof is based on Karagila's lecture note and Jech's textbook, which align well with the c.t.m.\ approach. 
However, some parts required adopting alternative proof methods.
The formalization process revealed insights into the challenges of formalizing set theory and the potential for future improvements.
